---
title: "EC 320 Problem Set 2"
date: "Winter 2022"
name: "PUT YOUR NAME HERE"
header-includes:
   - \usepackage{enumitem}
output: pdf_document
---

```{r include=FALSE}
library(pacman)
p_load(kableExtra, knitr)
# Knitr options
knitr::opts_chunk$set(
  echo=F,
  warning = F,
  message = F
)  
```
```{r include=FALSE}
library(knitr)
library(kableExtra)
# Knitr options
knitr::opts_chunk$set(
  echo=F,
  warning = F,
  message = F
)  
```

**INSTRUCTIONS:**

**There are four questions in total. Please answer them all and show the steps of how you derived your answer to receive the full credit**


# 1. Regression Logic (8 points)
\begin{enumerate}[label=\alph*)]
  \item TEXT.
  \item TEXT.
  \item TEXT.
  \item TEXT.
\end{enumerate}

\newpage

# 2. Calculating OLS estimates (12 points)
\begin{enumerate}[label=\alph*)]
  \item TEXT.
  \item TEXT.
  \item TEXT.
  \item TEXT.
\end{enumerate}

\newpage

# 3. Proof (8 points)
\begin{enumerate}[label=\alph*)]
  \item Residuals sum to zero
  \item The sample covariance between the independent variable and the residuals is zero.
\end{enumerate}

\newpage

# 4. Goodness of Fit (12 points)
\begin{enumerate}[label=\alph*)]
  \item Calculate the Total Sum of Squares
  \item Calculate the Residual Sum of Squares
  \item Calculate $R^2$
  \item Show that TSS=RSS+ESS.
\end{enumerate}




