---
title: "Midterm Review"
subtitle: "EC 320: Introduction to Econometrics"
date: "Winter 2022"
output:
  xaringan::moon_reader:
    css: ['default', 'metropolis', 'metropolis-fonts', 'my-css.css']
    # self_contained: true
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, middle

```{r Setup, include = F}
options(htmltools.dir.version = FALSE)
library(pacman)
# devtools::install_github("dill/emoGG")
library(pacman)
p_load(
  broom, here, tidyverse,
  emoGG, latex2exp, ggplot2, ggthemes, viridis, extrafont, gridExtra,
  kableExtra,
  data.table, dagitty, ggdag,
  dplyr, gganimate,
  lubridate,
  magrittr, knitr, parallel
)
# Define colors
red_pink <- "#e64173"
met_slate <- "#23373b" # metropolis font color
# Notes directory
dir_slides <- "~/GitHub/EC320_Econometrics/Lectures/06-Midterm_Review/"
# Knitr options
opts_chunk$set(
  comment = "#>",
  fig.align = "center",
  fig.height = 7,
  fig.width = 10.5,
  #dpi = 300,
  #cache = T,
  warning = F,
  message = F
)  
theme_simple <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 14),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank(),
  plot.title = element_blank(),
  legend.position = "none"
)
theme_gif <- theme_bw() + theme(
  axis.line = element_line(color = met_slate),
  panel.grid = element_blank(),
  rect = element_blank(),
  strip.text = element_blank(),
  text = element_text(family = "Fira Sans", color = met_slate, size = 14),
  axis.text.x = element_text(size = 12),
  axis.text.y = element_text(size = 12),
  axis.ticks = element_blank()
)
```


class: inverse, middle

# Practice

---
# Regression Table

## Example: 2016 election

**Q:** Write down the regression model estimated in column 1.

```{R, echo = F, escape = F}
tab <- data.frame(
  v1 = c("Intercept", "", "White (%)", "", "Poverty (%)", ""),
  v2 = rbind(
    c(-40.7, 0.910, ""),
    c("(1.95)", "(0.024)", "")
  ) %>% as.vector(),
  v3 = rbind(
    c(42.0, "", -0.647),
    c("(1.49)", "", "(0.087)")
  ) %>% as.vector(),
  v4 = rbind(
    c(-65.7, 1.05, 0.883),
    c("(2.99)", "(0.027)", "(0.081)")
  ) %>% as.vector()
) %>% kable(
  escape = F,
  col.names = c("Explanatory variable", "1", "2", "3"),
  align = c("l", rep("c", 4)),
  caption = "Outcome: Trump Margin (%)"
) %>%
row_spec(1:6, color = met_slate) %>%
row_spec(seq(2,6,2), color = "#c2bebe") %>%
row_spec(1:6, extra_css = "line-height: 110%;") %>%
column_spec(1, color = "black", italic = T)
tab %>% column_spec(2, bold = T)
```

--

**A:** $\text{Trump}_i = \beta_0 + \beta_1 \text{White}_i + u_i$.

---
# Regression Table

## Example: 2016 election

**Q:** Write down the regression model estimated in column 2.

```{R, echo = F}
tab %>% column_spec(3, bold = T)
```

--

**A:** $\text{Trump}_i = \beta_0 + \beta_1 \text{Poverty}_i + u_i$.

---
# Regression Table

## Example: 2016 election

**Q:** Write down the regression model estimated in column 3.

```{R, echo = F}
tab %>% column_spec(4, bold = T)
```

--

**A:** $\text{Trump}_i = \beta_0 + \beta_1 \text{White}_i + \beta_2\text{Poverty}_i + u_i$.



---
# Regression Table

## Example: 2016 election

**Q:** Does omitting $\text{White}_i$ bias the estimator of the $\text{Poverty}_i$ coefficient? 

```{R, echo = F}
tab %>% column_spec(c(3,4), bold = T)
```

--

**A:** *"Omitting* $\text{White}_i$ *appears to cause negative omitted-variable bias. The size of the bias is* $\beta_{\text{Poverty}}^{\text{Short}} - \beta_{\text{Poverty}}^{\text{Long}} = -0.647 - 0.883 = -1.53$*."*

---
# Omitted Variables

**Goal:** Isolate the effect of .hi[X] on .hi[Y]. <br> **Q:** Which variables, if omitted, would cause omitted-variable bias?

```{r, dev = "svg", echo = F, fig.height = 5}
dag1 <- dagify(Y ~ X + a + c + d,
               X ~ a + b + d,
               exposure = "X",
               outcome = "Y")

ggplot(data = dag1, aes(x = x, y = y, xend = xend, yend = yend)) +
      geom_dag_point(color = red_pink) +
      geom_dag_edges() +
      geom_dag_text(family = "Fira Sans") +
      theme_dag()
```

--

**A:** .pink[a] and .pink[d].




